Spatial Reasoning and Statistical Graphics

### 2014-10-19 Paper/Draft1/TestingVisualAptitude.Rnw
Justify use of the VST, addition of references.

\section{Methods}
\comment{Besides pointing out what we asked participants to do, we also need some why, and also why in particular these tests were selected. part of the validity consideration can go in here - and the classification from the naval test of the tests we picked into different classes (maybe already with a hint that for our data these weren't actually orthogonal classes, because we found a really high correlation between card rotation and paper folding.)\\
\textit{I don't know if the tests are supposed to be orthogonal. It would be hard to construct a set of tests that was orthogonal, because the abilities themselves aren't easily separable - math reasoning and spatial reasoning are related, for instance, and verbal abilities are correlated with creativity too.}}
Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once testing is complete). Tasks are designed so that participants are under time pressure; they are not expected to complete all of the problems in each section. This provides more discrimination between high scorers and prevents score compression at the top of the range. 
\begin{itemize}
\item Visual Search Task: designed to test participants' ability to find a target stimulus in a field of distractors. An example is shown in figure \ref{fig:VST}. 
\begin{figure}[htbp]\centering
\includegraphics[width=.7\textwidth]{VisualSearch}
\caption[Visual Search Task]{Visual Search Task. Participants are instructed to find the plot numbered 1-24 which matches the plot labeled "Target". Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
\end{figure}
\item Paper Folding Task: tests participants' ability to visualize and mentally manipulate figures in three dimensions. Associated with the ability to extrapolate symmetry and reflection over multiple steps. An example is shown in figure \ref{fig:paperfolding}.
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{paperfolding}
\caption[Paper Folding Task]{Paper Folding Task. Participants are instructed to pick the figure matching the sequence of steps shown in the left-hand figure. Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
\end{figure}
\item Card Rotation Task: tests participant's ability to rotate objects in two dimensions to distinguish between left-hand and right-hand versions of the same figure. Tests spatial reasoning ability and mental rotation skills. An example is shown in figure \ref{fig:cardrotation}.
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{cardrotation}
\caption[Card Rotation Task]{Card Rotation Task. Participants mark each figure on the right hand side as either the same or different than the figure on the left hand side of the dividing line. Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
\end{figure}
\item Figure Classification Task: tests participant's ability to extrapolate rules from provided figures. This task is associated with visual reasoning capabilities and we expect that it should correlate with the ability to pick out a signal plot from a lineup. An example is shown in figure \ref{fig:figureclassification}. 
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{figureclassification}
\caption[Figure Classification Task]{Figure Classification Task. Participants classify each figure in the second row as belonging to group 1, 2, or 3 (if applicable). Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.}\label{fig:figureclassification}
\end{figure}
\end{itemize}

>>>>

\section{Methods}
\comment{Besides pointing out what we asked participants to do, we also need some why, and also why in particular these tests were selected. part of the validity consideration can go in here - and the classification from the naval test of the tests we picked into different classes (maybe already with a hint that for our data these weren't actually orthogonal classes, because we found a really high correlation between card rotation and paper folding.)\\
\textit{I don't know if the tests are supposed to be orthogonal. It would be hard to construct a set of tests that was orthogonal, because the abilities themselves aren't easily separable - math reasoning and spatial reasoning are related, for instance, and verbal abilities are correlated with creativity too.}}
Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once testing is complete). Tasks are designed so that participants are under time pressure; they are not expected to complete all of the problems in each section. This provides more discrimination between high scorers and prevents score compression at the top of the range. 
\begin{itemize}
\item Visual Search Task: designed to test participants' ability to find a target stimulus in a field of distractors. An example is shown in figure \ref{fig:VST}. The visual search task is similar in concept to lineups: it tests one's ability to find the target plot. Historically, it has been used as a measure of brain damage\citep{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in \citet{anderson1983interactive}). The similarity to lineup protocol as well as the simiplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations. 
\begin{figure}[htbp]\centering
\includegraphics[width=.7\textwidth]{VisualSearch}
\caption[Visual Search Task]{Visual Search Task. Participants are instructed to find the plot numbered 1-24 which matches the plot labeled "Target". Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
\end{figure}
\item Paper Folding Task: tests participants' ability to visualize and mentally manipulate figures in three dimensions. Associated with the ability to extrapolate symmetry and reflection over multiple steps. An example is shown in figure \ref{fig:paperfolding}.
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{paperfolding}
\caption[Paper Folding Task]{Paper Folding Task. Participants are instructed to pick the figure matching the sequence of steps shown in the left-hand figure. Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
\end{figure}
\item Card Rotation Task: tests participant's ability to rotate objects in two dimensions to distinguish between left-hand and right-hand versions of the same figure. Tests spatial reasoning ability and mental rotation skills. An example is shown in figure \ref{fig:cardrotation}.
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{cardrotation}
\caption[Card Rotation Task]{Card Rotation Task. Participants mark each figure on the right hand side as either the same or different than the figure on the left hand side of the dividing line. Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
\end{figure}
\item Figure Classification Task: tests participant's ability to extrapolate rules from provided figures. This task is associated with visual reasoning capabilities and we expect that it should correlate with the ability to pick out a signal plot from a lineup. An example is shown in figure \ref{fig:figureclassification}. 
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{figureclassification}
\caption[Figure Classification Task]{Figure Classification Task. Participants classify each figure in the second row as belonging to group 1, 2, or 3 (if applicable). Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.}\label{fig:figureclassification}
\end{figure}
\end{itemize}
